---
title: "Assignament Maching Learning"
author: "WIlmer Fuentes Neira"
date: "Sunday, February 22, 2015"
output:
  html_document:
    fig_height: 6
    fig_width: 9
---
markdownToHTML("final_job.md", "final_job.html")

## Preparando los dato

```{r}
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)

```


```{r}
options(warn=-1)
library(caret)
library(randomForest)
library(Hmisc)
library(foreach)
library(doParallel)
set.seed(1235)
```
## Lectura de Datos cargados
```{r, echo=F, result="hide"}
training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

```{r,echo=F, result="hide"}
#summary(training)
#describe(training)
#sapply(data, class)
#sapply(training, class)
#sapply(testing, class)

#str(data)
```

```

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
witraining <- training[inTrain, ]; myTesting <- training[-inTrain, ]
dim(witraining); dim(myTesting)

##Cleaning

metricdata <- nearZeroVar(witraining, saveMetrics=TRUE)
metricvars <- names(witraining) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
witraining <- witraining[!metricvars]
#To check the new N?? of observations
dim(witraining)
```


Transformation 2: Killing first column of Dataset - ID Removing first ID variable so that it does not interfer with ML Algorithms:


```{r}
witraining <- witraining[c(-1)]
modFitA1 <- rpart(classe ~ ., data=witraining, method="class")

```

Transformation 3: Cleaning Variables with too many NAs. For Variables that have more than a 60% threshold of NA's I'm going to leave them out:

```{r}
trainingV3 <- witraining #creating another subset to iterate in loop
for(i in 1:length(witraining)) { #for every column in the training dataset
        if( sum( is.na( witraining[, i] ) ) /nrow(witraining) >= .6 ) { #if n?? NAs > 60% of total observations
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(witraining[i]), names(trainingV3)[j]) ) ==1)  { #if the columns are the same:
                trainingV3 <- trainingV3[ , -j] #Remove that column
            }   
        } 
    }
}
```


```{r}
#Seting back to our set:
witraining <- trainingV3
rm(trainingV3)
```

Now let us do the exact same 3 transformations but for our myTesting and testing data sets.

```{r}
clean1 <- colnames(witraining)
clean2 <- colnames(witraining[, -58]) #already with classe column removed
myTesting <- myTesting[clean1]
testing <- testing[clean2]

#To check the new N?? of observations
dim(myTesting)
```


```{r}

dim(testing)
```



```{r}
for (i in 1:length(testing) ) {
        for(j in 1:length(witraining)) {
        if( length( grep(names(witraining[i]), names(testing)[j]) ) ==1)  {
            class(testing[j]) <- class(witraining[i])
        }      
    }      
}
#And to make sure Coertion really worked, simple smart ass technique:
testing <- rbind(witraining[2, -58] , testing) #note row 2 does not mean anything, this will be removed right.. now:
testing <- testing[-1,]
```




```{r }
fancywi<-fancyRpartPlot(modFitA1)

```

Predicting:

```{r}
predictionsA1 <- predict(modFitA1, myTesting, type = "class", na.action = na.omit)
```


```{r}
confusionMatrix(predictionsA1, myTesting$classe)
```
##Using ML algorithms for prediction: Random Forests

```{r}
modFitB1 <- randomForest(classe ~. , data=witraining, na.action = na.omit)
```


Predicting in-sample error:
```{r}
predictionsB1 <- predict(modFitB1, myTesting, type = "class")
```


(Moment of truth) Using confusion Matrix to test results:

```{r}
confusionMatrix(predictionsB1, myTesting$classe)
```

Random Forests yielded better Results, as expected!

## Generating Files to submit as answers for the Assignment:

Finally, using the provided Test Set out-of-sample error.

For Random Forests we use the following formula, which yielded a much better prediction in in-sample:

```{r}
predictionsB2 <- predict(modFitB1, testing, type = "class")
```

```{r}
pml_write_files = function(x){
  n = length(x)
 for(i in 1:n){
   filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionsB2)
```




